The Space of Feasible Execution Times for Asynchronous Periodic Task Systems using Definitive Idle Times

### 2013-09-29 dit-paper_ieee.tex
multi A crp description rewrite

	\subsection{Solving the modular system}
		A sketch of the construction algorithm to compute $t_d$ can be found in
		Algorithm~\ref{alg:algoCRP} (note that $x^{-1} \; (mod \; y)$ means the
		modular inverse of $x$ in the naturals modulo $y$). The congruences must
		first be modified in order to obtain pairwise coprime moduli. Each $n_i$
		is factored into prime powers $p^{k_1}_1 p^{k_2}_2 \cdots p^{k_m}_m = n_i$
		with $k_j$ the highest integer s.t. $p^{k_j}_j$ divides $n_i \; \forall j
		\; \mid 1 \leqslant j \leqslant m$. Each original constraint $t_d \equiv x_i
		\; (mod \; n_i)$ is replaced by the equivalent set \cite{knuth1969art}
		\begin{equation*}
			\begin{split}
				t_d \equiv x_i 	& \; (mod \; p^{k_1}_1) \\
				t_d \equiv x_i 	& \; (mod \; p^{k_2}_2) \\
								& \vdots \\
				t_d \equiv x_i 	& \; (mod \; p^{k_m}_m) \\
			\end{split}
		\end{equation*}
		This operation (later called constraint \emph{factorisation}) might create
		duplicate constraints, which are removed.
		As shown in \cite{knuth1969art}, the resulting system can be solved using
		Algorithm~\ref{alg:algoCRP}. For ease of reading, we call $C$ the set of
		resulting constraints $c$ each written $t_d \equiv x_c \; (mod \;
		p_c)$.

		\begin{algorithm}
			\caption{Gauss's CRP Algorithm}
			\label{alg:algoCRP}
			\begin{algorithmic}[1]
				\REQUIRE $H$ is the hyperperiod of the system
				\FORALL{$c \in C$}
					\STATE $Mc \leftarrow \frac{H}{p_c}$
					\STATE $invMc \leftarrow Mc^{-1} \; (mod \; p_c)$
					\STATE $e_c \leftarrow Mc \cdot invMc$
				\ENDFOR
				\RETURN $\sum\limits_{c \in C}{x_c \cdot e_c} \; (mod \; H)$
			\end{algorithmic}
		\end{algorithm}

		Algorithm~\ref{alg:algoCRP} finds one DIT given one combination of $a_i$
		values. However using it in practice is very inefficient as there are
		exponentially many different combinations of $a_i$ values in the number of
		tasks (the cartesian product of all the $A_i$ sets), and resolving
		exponentially many CRP instances is prohibitively slow.
		In this paper a faster method for finding the FPDIT is proposed. It consists
		in merging all these CRP instances into one problem (finding all DITs), which can be solved faster by
		Algorithm~\ref{alg:multiACRP}, as a significant part of the intermediate
		results are shared between instances. The first part consisting in splitting
		all constraints into sets of constraints with prime power moduli is slightly
		different. We define for each task $\tau_i$ and $a_j \in A_i$ :
		$D_{ij}$ the set of constraints $c$ of the form $t_d \equiv x_c
		\; (mod \; p_c)$ resulting from the factorisation of $t_d
		\equiv a_j+O_i \; (mod \; n_i)$. We define $C = \bigcup\limits_{\tau_i \in \tau,
		a_j \in A_i}{C_{ij}}$, $P$ the set of all $p_c$ values appearing in
		constraints of $C$ and $C_{p}$ the subset of $C$ containing all constraints
		$c$ with $p_c = p$.

		\begin{algorithm}
			\caption{Modified Gauss Algorithm}
			\label{alg:multiACRP}
			\begin{algorithmic}[1]
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE $sumchunks \leftarrow \text{empty map}$
				\FORALL{$p \in P$}
					\STATE $taskTerms \leftarrow \text{empty list}$
					\STATE $Mp \leftarrow \frac{H}{p}$
					\STATE $invMp \leftarrow Mp^{-1} (mod \; p)$
					\FORALL{$c \in C_p$}
						\STATE append $(x_c \cdot Mp \cdot invMp)$ to $taskTerms$
					\ENDFOR
					\STATE $sumchunks[p] \leftarrow taskTerms$
				\ENDFOR

				\STATE $results \leftarrow \text{empty list}$
				\STATE append $0$ to $results$
				\FORALL{$p \in P$}
					\STATE $newresults \leftarrow \text{empty list}$
					\FORALL{$chunk \in sumchunks[p]$}
						\FORALL{$r \in results$}
							\STATE $\text{append} \; (r + chunk) \; \text{to} \; newresults$
						\ENDFOR
					\ENDFOR
					\STATE $results \leftarrow newresults$
				\ENDFOR
				\RETURN $r \; (mod \; H)$ for all $r \in results$
			\end{algorithmic}
		\end{algorithm}

		This method ``forgets'' the link between generated constraints and the tasks
		(with the associated $A_i$) they come from. Constraints are instead grouped by their
		primal power modulo $p$ and each $p$ is associated with a set of $x_c$ values
		for $c$ in $C_p$, acting similarly as the $A_i$ set of each task. Due to this
		simplification, Algorithm~\ref{alg:multiACRP} might output ``false''
		solutions. This isn't a major drawback as checking whether a solution is
		correct can be done in linear time in the number of tasks (for each task
		$\tau_i$, check whether at least one $a \in A_i$ satisfies $t_d \equiv a \;
		(mod \; T_i)$) and the only interesting result is the minimum $t_d$ which
		means that using a minimum heap structure containing the results we only need
		to check solutions for correctness until the first (hence minimum) correct
		solution is found.

		Both Algorithm~\ref{alg:algoCRP} and Algorithm~\ref{alg:multiACRP} require the
		computation of one modular inversion per constraint and per $a_i$ value, which
		is done in a linear number of modular inversions per task. However, to
		find all DITs Algorithm~\ref{alg:algoCRP} has to be executed once for each possible
		combination of $a_i$ values, which is exponential in the number of tasks.
		Algorithm~\ref{alg:multiACRP} has the same theoretical complexity as the
		computation of all DITs still requires an exponential number of answers in the
		number of tasks, but it is nonetheless significantly faster in practice. The
		running time of Algorithm~\ref{alg:multiACRP} is also less sensitive to the
		size of the $A_i$ sets per task, which means that this method
		remains practical when tasks deadlines are much smaller than periods.

>>>>

	\subsection{Solving the modular system}
		A sketch of the construction algorithm to compute $t_d$ can be found in
		Algorithm~\ref{alg:algoCRP} (note that $x^{-1} \; (mod \; y)$ means the
		modular inverse of $x$ in the naturals modulo $y$). The congruences must
		first be modified in order to obtain pairwise coprime moduli. Each $n_i$
		is factored into prime powers $p^{k_1}_1 p^{k_2}_2 \cdots p^{k_m}_m$
		with $k_j$ the highest integer s.t. $p^{k_j}_j$ divides $n_i \; \forall j
		\; \mid 1 \leqslant j \leqslant m$. Each original constraint $t_d \equiv x_i
		\; (mod \; n_i)$ is replaced by the equivalent set \cite{knuth1969art}
		\begin{equation*}
			\begin{split}
				t_d \equiv x_i 	& \; (mod \; p^{k_1}_1) \\
				t_d \equiv x_i 	& \; (mod \; p^{k_2}_2) \\
								& \vdots \\
				t_d \equiv x_i 	& \; (mod \; p^{k_m}_m) \\
			\end{split}
		\end{equation*}
		This operation might create
		duplicate constraints, which are removed.
		As shown in \cite{knuth1969art}, the resulting system can be solved using
		Algorithm~\ref{alg:algoCRP}. For ease of reading, we call $C$ the set of
		resulting constraints. We define two functions on the set of
		constraints. Assuming contraint $c$ of the form $t_d \equiv x \; (mod \; p)$,
		we have $p(c) = p$ and $x(c) = x$

		\begin{algorithm}
			\caption{Gauss's CRP Algorithm}
			\label{alg:algoCRP}
			\begin{algorithmic}[1]
				\REQUIRE $H$ is the hyperperiod of the system
				\FORALL{$c \in C$}
					\STATE $Mc \leftarrow \frac{H}{p(c)}$
					\STATE $invMc \leftarrow Mc^{-1} \; (mod \; p(c))$
					\STATE $e_c \leftarrow Mc \cdot invMc$
				\ENDFOR
				\RETURN $\sum\limits_{c \in C}{x(c) \cdot e_c} \; (mod \; H)$
			\end{algorithmic}
		\end{algorithm}

		Algorithm~\ref{alg:algoCRP} finds one DIT given one combination of $a_i$
		values. However using it in practice is very inefficient as there are
		exponentially many different combinations of $a_i$ values in the number of
		tasks (the cartesian product of all the $A_i$ sets), and resolving
		exponentially many CRP instances is prohibitively slow.
		In this paper a faster method for finding the FPDIT is proposed. It consists
		in merging all these CRP instances into one problem (finding all DITs), which can be solved faster by
		Algorithm~\ref{alg:multiACRP}, as a significant part of the intermediate
		results are shared between instances. First, the same $C$ set is obtained,
		but instead of one $x_i$ value per task, all variants in $X_i$ spawn a set of
		constraints. All the duplicate and useless constraints are removed. Then we
		define $P$ the set of all powers of prime numbers appearing as moduli in the
		congruences (the $p_j^{k_j}$ values). For all $p \in P$, we define $C_p$ the set of
		constraints where $p$ is the modulo.
		\begin{algorithm}
			\caption{Modified Gauss Algorithm}
			\label{alg:multiACRP}
			\begin{algorithmic}[1]
                \REQUIRE $P$, $C_{p}$, $x(c)$: as defined above
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE $sumchunks \leftarrow \text{empty mapping}$
				\FORALL{$p \in P$}
					\STATE $taskTerms \leftarrow \text{empty list}$
					\STATE $Mp \leftarrow \frac{H}{p}$
					\STATE $invMp \leftarrow Mp^{-1} (mod \; p)$
					\FORALL{$c \in C_p$}
						\STATE append $(x(c) \cdot Mp \cdot invMp)$ to $taskTerms$
					\ENDFOR
					\STATE $sumchunks[p] \leftarrow taskTerms$
				\ENDFOR

				\STATE $results \leftarrow \text{empty list}$
				\STATE append $0$ to $results$
				\FORALL{$p \in P$}
					\STATE $newresults \leftarrow \text{empty list}$
					\FORALL{$chunk \in sumchunks[p]$}
						\FORALL{$r \in results$}
							\STATE $\text{append} \; (r + chunk) \; \text{to} \; newresults$
						\ENDFOR
					\ENDFOR
					\STATE $results \leftarrow newresults$
				\ENDFOR
				\RETURN $r \; (mod \; H)$ for all $r \in results$
			\end{algorithmic}
		\end{algorithm}
		In the first loop over $P$ of Algorithm~\ref{alg:multiACRP}, all the
		required intermediate values are computed. In the second loop, those
		intermediate values are added in every possible way. Both
		Algorithm~\ref{alg:algoCRP} and Algorithm~\ref{alg:multiACRP} require the
		computation of one modular inversion per constraint, which
		is done in a linear number of modular inversions per task. However, to
		find all DITs Algorithm~\ref{alg:algoCRP} has to be executed once for each possible
		combination of $a_i$ values, which is exponential in the number of tasks.
		Algorithm~\ref{alg:multiACRP} has the same theoretical complexity as the
		computation of all DITs still requires an exponential number of answers in the
		number of tasks, but it is nonetheless significantly faster in practice. One
		drawback of Algorithm~\ref{alg:multiACRP} is that the replacement of each
		constraint by sets of constraints with powers of primes as moduli is not
		equivalent. Algorithm~\ref{alg:multiACRP} might ouput values that are not
		DITs. However, as we only need the FPDIT, it remains practical to check those
		solutions one by one from the lowest to the highest, and stop as soon as the
		first correct result is found.
